---
title: "Residential Property Features and Land Use Data"
author: "Leanne Chan"
date: "3/23/2020"
output: html_document
---

# Set up 
```{r setup, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
library(tigris)
options(tigris_use_cache = TRUE) 
library(tidycensus)
library(viridis)
library(riem)                
library(gridExtra)
library(knitr)
library(kableExtra)
library(RSocrata)
library(gifski)
library(png)
library(ggsn)
library(gganimate)
library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(car) #test colinarity
library(ggplot2)
library(dplyr)
library(gmodels)
library(MASS)
library(e1071)
library(Hmisc)
library(Boruta)
library(corrplot)
library(viridis)
library(stargazer)
library(tigris)
library(GoodmanKruskal)
library(spatstat)
library(riem)
library(readr)
library(readxl)
library(QuantPsyc)
library(lubridate)
library(jsonlite)
library(plyr)
library(ggpmisc)

```

```{r map theme}
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2)
  )
}

qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                          c(.01,.2,.4,.6,.8), na.rm=T)
  }
}

q5 <- function(variable) {as.factor(ntile(variable, 5))}
palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")

```

```{r philly and fire data, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
# city limits
philly <- 
  st_read("http://data.phl.opendata.arcgis.com/datasets/405ec3da942d4e20869d4e1449a2be48_0.geojson")%>%st_transform(crs=2272)

st_crs(philly)


fishnet <- 
  st_make_grid(philly, cellsize = 800) %>%
  st_sf()

fishnet <- 
  fishnet[philly,] %>%
  mutate(uniqueID = rownames(.)) %>%
  dplyr::select(uniqueID)


#### Fire data
fire_excel<-read_excel('C:/Users/leann/OneDrive/Desktop/MUSA 800/Data/2015 to 2019 fires for u of pa report run 1620.xlsx')
fire.sf <- 
  fire_excel%>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, agr = "constant") %>%
  st_set_crs(4326) %>%
  st_transform(2272)
fire.sf <- 
  st_intersection(fire.sf, philly)
fire.sf <- 
  fire.sf %>%
  mutate(date=dmy(alm_date),
         Year=year(date),
         Month=month(date))

## kernel density of fires by year 
kernelDensity <- function(pointData, year, fishnetGrid, name){
  data<-pointData %>%filter(Year == year)
  count<- data %>% #get count by fishnet
  dplyr::select() %>% 
  mutate(count = 1) %>% 
  aggregate(., fishnetGrid, sum) %>%
  mutate(count = ifelse(is.na(count), 0, count),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnetGrid) / 24), size=nrow(fishnetGrid), replace = TRUE))
  
  point_ppp <- as.ppp(st_coordinates(data), W = st_bbox(count))
  KD <- spatstat::density.ppp(point_ppp, 1000)
  
  KD_dataframe<-as.data.frame(KD) %>% 
    st_as_sf(coords = c("x", "y"), crs = st_crs(fishnetGrid)) %>%
    aggregate(.,fishnetGrid, mean)%>%plyr::rename(.,c('value' = name))
    
  return(KD_dataframe)
}

f18_KD <- kernelDensity(fire.sf, 2018, fishnet, 'Fire18KD')
f19_KD <- kernelDensity(fire.sf, 2019, fishnet, 'Fire19KD')
```

# Residential Building Data
```{r get OPA data residential only}
raw<-fromJSON('https://phl.carto.com/api/v2/sql?q=SELECT%20census_tract,%20category_code,%20market_value,%20%20parcel_number,%20sale_price,%20total_area,%20total_livable_area,%20year_built,%20zip_code,%20ST_Y(the_geom)%20AS%20lat,%20ST_X(the_geom)%20AS%20lng%20FROM%20opa_properties_public%20WHERE%20category_code%20=%20%271%27')
formated<- as.data.frame(raw$rows)

# sum(is.na(formated$sale_price)) # none missing but a lot of 1s. 
# sum(is.na(formated$market_value)) #290 missing
# sum(is.na(formated$year_built)) # 2 missing
# sum(is.na(formated$total_livable_area)) # 2 missing
# sum(is.na(formated$total_area))
# sum(is.na(formated$census_tract))
# sum(is.na(formated$parcel_number))
# sum(is.na(formated$zip_code))
# sum(formated$sale_price==1) #122394
# sum(formated$market_value==1) #0
# sum(formated$market_value==0) #69
# sum(formated$total_livable_area==0) #463
# sum(formated$total_area==0) # 33609
formated <- na.omit(formated)

final_data <-formated%>%
   dplyr::filter(formated$total_livable_area != 0 & formated$market_value != 0 & formated$year_built!='0000')%>%
  dplyr::select(market_value, total_livable_area, year_built, census_tract, zip_code, parcel_number, lat , lng) %>%na.omit(., cols=c("lng", "lat"))%>%
    st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs=2272)%>%
    mutate(
      X = st_coordinates(.)[,1],
      Y = st_coordinates(.)[,2],
      val_per_size = market_value/total_livable_area,
      year_built = as.numeric(year_built),
    )

# more mutate for the year built
final_data<- final_data%>%mutate(year_cat=  cut2(final_data$year_built, c(1650, 1700, 1750, 1800, 1850, 1900, 1950, 2000, 2020)), building_age = 2020-year_built)
```

```{r house value data}
values_plot<-ggplot() + 
  geom_sf(data=philly)+
  geom_sf(data = final_data, aes(colour = q5(val_per_size)), show.legend = "point", size = .5) +
  scale_colour_viridis_d(option='inferno',
                         direction=-1,
                      labels=qBr(final_data,"val_per_size"),
                      name="Quintile\nBreaks") +
  labs(title = "House Values in Philadelphia",
       subtitle='House Values= market value/total livable area.\nData from Office of Property Assesments.') +
  mapTheme()

values_plot

# # continous scale -- NOPE, its too skewed.
# values_plot_C<-ggplot() + 
#   geom_sf(data=philly)+
#   geom_sf(data = final_data, aes(color=val_per_size), show.legend = "point", size = .5) +
#   scale_colour_viridis(option='inferno',
#                          direction=-1,
#                       name="Continuous Scale") +
#   labs(title = "House Values in Philadelphia", subtitle='House Values= market value/total livable area.\nData from Office of Property Assesments.') +
#   mapTheme()
# 
# values_plot_C
```

```{r year built data}
#built_year discrete
year_plot<-ggplot() + 
  geom_sf(data=philly)+
  geom_sf(data = final_data, aes(colour = final_data$year_cat), show.legend = "point", size = .5) +
  scale_colour_viridis_d(option = 'inferno', name="50 Year Breaks") +
  labs(title = "Year Built of Residential Houses in Philadelphia", subtitle='Data from Office of Property Assesments.') + mapTheme()

year_plot

# age instead continous
age_plot<-ggplot() + 
  geom_sf(data=philly)+
  geom_sf(data = final_data, aes(colour = final_data$building_age), show.legend = "point", size = .5) +
  scale_colour_viridis(option = 'inferno', name="Building Age", direction=-1) +
  labs(title = "Age of Residential Houses in Philadelphia", subtitle='Data from Office of Property Assesments.') + mapTheme()

age_plot

```

```{r aggregte to fishnet}
# Mean value of houses and age of houses per fishnet 
fishnet_meanVal<-final_data %>% 
    dplyr::select('val_per_size') %>% 
    mutate(count = 1) %>% 
    aggregate(., fishnet, mean)%>%
    mutate(count = ifelse(is.na(count), 0, count),
           uniqueID = rownames(.),
           cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE))%>%
  mutate(val_per_size=replace_na(val_per_size,0))

fishnet_medianAge<-final_data %>% 
    dplyr::select('building_age') %>% 
    mutate(count = 1) %>% 
    aggregate(., fishnet, median)%>%
    mutate(count = ifelse(is.na(count), 0, count),
           uniqueID = rownames(.),
           cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE))%>%
  mutate(building_age=replace_na(building_age,0))

#### PLOTS 
## Value plot 

houseVal_fishnet_plot<-ggplot() + 
  geom_sf(data=philly)+
  geom_sf(data = fishnet_meanVal, aes(colour=val_per_size)) +
  scale_colour_viridis(option='inferno',
                         direction=-1,
                      name="Average Value Per Size Houses") +
  labs(title = "Average House Values in Philadelphia", subtitle='Values are aggregated by 800ft grid cells.\nData from Office of Property Assesments.') +
  mapTheme()

houseVal_fishnet_plot

### Age plot 
houseAge_fishnet_plot<-ggplot() + 
  geom_sf(data=philly)+
  geom_sf(data = fishnet_medianAge, aes(colour=building_age)) +
  scale_colour_viridis(option='inferno',
                         direction=-1,
                      name="Median Age of Houses") +
  labs(title = "Median House Age in Philadelphia", subtitle='Ages are aggregated by 800ft grid cells.\nData from Office of Property Assesments.') +
  mapTheme()

houseAge_fishnet_plot

```

The code here joins the aggregated Mean House Values and Median age to the fishnet. 
```{r aggregate to fire}

fishnet_final_data<-st_join(fishnet_meanVal, fishnet_medianAge %>% dplyr::select('building_age') %>% st_centroid(.))

fishnet_fireKD<-st_join(f19_KD, f18_KD %>% st_centroid(.))

fishnet_data_fire<-st_join(fishnet_final_data, fishnet_fireKD %>% st_centroid(.))

```

```{r plot with fire, fig.width=12, fig.height=10, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE,fig.align='center'}
#### SCATTER PLOTS ### 
houseData_2018<-fishnet_data_fire%>%
  dplyr::select('Fire18KD', 'building_age', 'val_per_size')%>%
  st_set_geometry(NULL)%>%
  gather(Variable, Value, -Fire18KD)%>%
   ggplot(aes(Value, Fire18KD)) +
     geom_point() +
     geom_smooth(method = "lm", se=F, colour = "#FA7800") +
    stat_poly_eq(formula = y ~ x,
               eq.with.lhs = "italic(hat(y))~`=`~",
               aes(label = paste(..rr.label.., sep =" ")), 
               parse = TRUE, label.x = 60, label.y = 70) +    
     facet_wrap(~Variable, ncol = 2, scales = "free") + labs(title = 'Residential Property Data',subtitle = '2018 Fire Data')


houseData_2019<-fishnet_data_fire%>%
  dplyr::select('Fire19KD', 'building_age', 'val_per_size')%>%
  st_set_geometry(NULL)%>%
  gather(Variable, Value, -Fire19KD)%>%
   ggplot(aes(Value, Fire19KD)) +
     geom_point() +
     geom_smooth(method = "lm", se=F, colour = "#FA7800") +
    stat_poly_eq(formula = y ~ x,
               eq.with.lhs = "italic(hat(y))~`=`~",
               aes(label = paste(..rr.label.., sep =" ")), 
               parse = TRUE, label.x = 60, label.y = 70) +    
     facet_wrap(~Variable, ncol = 2, scales = "free") + labs(subtitle = '2019 Fire Data')



grid.arrange(houseData_2018, houseData_2019, ncol=1, top="Kernel Density of Fires to Property Data*;\n Source: OPA Violations", bottom='*: Property Data is from past 5 years as assessment date data was unavailable.')


```

# Land use Data 
```{r land use data}
landUse<-st_read('C:/Users/leann/OneDrive/Desktop/MUSA 800/Data/Land_Use')%>%st_transform(crs=2272)

landUse<-landUse%>%mutate(C_DIG1DESC = as.factor(C_DIG1DESC),
                                  C_DIG1 = as.numeric(C_DIG1))%>%
  mutate(is_residential = case_when(
    C_DIG1==1 ~ 1,
    C_DIG1!=1 ~ 0
  ),
  is_commercial=case_when(
    C_DIG1==2~1,
    C_DIG1!=2~0
  ),
  is_industrial = case_when(
    C_DIG1==3 ~1,
    C_DIG1!=3~0,
  ))
  
class(landUse$is_residential)
summary(landUse)
sum(landUse$C_DIG1==1)

```

```{r land use plot, fig.width=12, fig.height=10, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE,fig.align='center'}

landuseforplot<-landUse%>%mutate(is_residential=as.factor(is_residential),
                                 is_commercial=as.factor(is_commercial),
                                 is_industrial=as.factor(is_industrial))


residential_plot<-ggplot() + 
  geom_sf(data=philly)+
  geom_sf(data = landuseforplot, aes(colour=landuseforplot$is_residential)) +
  scale_colour_viridis_d(option = 'inferno', name="Residential", labels=c('No', 'Yes')) +
  labs(title = "Land Uses in Philadelphia", subtitle='Data from Philadelphia Department of Planning and Development.') + mapTheme()


commercial_plot<-ggplot() + 
  geom_sf(data=philly)+
  geom_sf(data = landuseforplot, aes(colour = landuseforplot$is_commercial)) +
  scale_colour_viridis_d(option = 'inferno', name="Commercial", labels=c('No', 'Yes')) +
  labs(title = "Land Uses in Philadelphia", subtitle='Data from Philadelphia Department of Planning and Development.') + mapTheme()


industrial_plot<-ggplot() + 
  geom_sf(data=philly)+
  geom_sf(data = landuseforplot, aes(colour = landuseforplot$is_industrial)) +
  scale_colour_viridis_d(option = 'inferno', name="Industrial", labels=c('No', 'Yes')) +
  labs(title = "Land Uses in Philadelphia", subtitle='Data from Philadelphia Department of Planning and Development.') + mapTheme()


grid.arrange(residential_plot, commercial_plot, industrial_plot, ncol=3, top="Land Uses in Philadelphia")

```

```{r landuse fishnet}
# since if parcel is residential is 1, the mean is the percent
fishnet_residential<-landUse %>% 
    dplyr::select(is_residential) %>%
    mutate(count = 1) %>%
    aggregate(., fishnet, mean)%>%
    mutate(count = ifelse(is.na(count), 0, count),
           uniqueID = rownames(.),
           cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE),
           percent_residential = is_residential)

fishnet_commercial<-landUse %>% 
    dplyr::select(is_commercial) %>% 
    mutate(count = 1) %>% 
    aggregate(., fishnet, mean)%>%
    mutate(count = ifelse(is.na(count), 0, count),
           uniqueID = rownames(.),
           cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE),
           percent_commercial = is_commercial)

fishnet_industrial<-landUse %>% 
    dplyr::select(is_industrial) %>% 
    mutate(count = 1) %>% 
    aggregate(., fishnet, mean)%>%
    mutate(count = ifelse(is.na(count), 0, count),
           uniqueID = rownames(.),
           cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE),
           percent_industrial = is_industrial)

```

```{r aggregate landuse to fishnet}
# add percent residential and percent industrial
fishnet_data_fire<-st_join(fishnet_data_fire, fishnet_residential 
                             %>% dplyr::select('percent_residential')
                             %>% st_centroid(.)%>%st_join(., fishnet_industrial
          %>% dplyr::select('percent_industrial')
          %>% st_centroid(.)%>%st_join(., fishnet_commercial
          %>% dplyr::select('percent_commercial')
          %>% st_centroid(.))))

```

```{r landuse plot with fire, fig.width=12, fig.height=10, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE,fig.align='center'}
#### SCATTER PLOTS ### 
landuse_2018<-fishnet_data_fire%>%
  dplyr::select('Fire18KD', 'percent_residential', 'percent_commercial','percent_industrial')%>%
  st_set_geometry(NULL)%>%
  gather(Variable, Value, -Fire18KD)%>%
   ggplot(aes(Value, Fire18KD)) +
     geom_point() +
     geom_smooth(method = "lm", se=F, colour = "#FA7800") +
    stat_poly_eq(formula = y ~ x,
               eq.with.lhs = "italic(hat(y))~`=`~",
               aes(label = paste(..rr.label.., sep =" ")), 
               parse = TRUE, label.x = 60, label.y = 70) +    
     facet_wrap(~Variable, ncol = 3, scales = "free") + labs(title = 'Land Use Data',subtitle = '2018 Fire Data')




landuse_2019<-fishnet_data_fire%>%
  dplyr::select('Fire19KD', 'percent_residential', 'percent_commercial','percent_industrial')%>%
  st_set_geometry(NULL)%>%
  gather(Variable, Value, -Fire19KD)%>%
   ggplot(aes(Value, Fire19KD)) +
     geom_point() +
     geom_smooth(method = "lm", se=F, colour = "#FA7800") +
    stat_poly_eq(formula = y ~ x,
               eq.with.lhs = "italic(hat(y))~`=`~",
               aes(label = paste(..rr.label.., sep =" ")), 
               parse = TRUE, label.x = 60, label.y = 70) +    
     facet_wrap(~Variable, ncol = 3, scales = "free") + labs(title = 'Land Use Data',subtitle = '2019 Fire Data')



grid.arrange(landuse_2018, landuse_2019, ncol=1, top="Kernel Density of Fires to Land Use Data\n Source:  Philadelphia Department of Planning and Development")


```

### The final fishnet with the 5 features is called the fishnet_data_fire and the features are: 
1. Median Age of Residential Buildings (building_age) 
2. Average value per size of Residential Buildings (val_per_size)
3. Percentage of Residential Land (percent_residential)
4. Percentage of Commercial Land (percent_commercial)
5. Percentage of Industrial Land (percent_industrial)

